We would like to further examine the top scorers, given the high number of records in the data observed in Section \ref{examine-further}. In Taiwan's education system, the top tier of high schools and colleges can be further segmented. The top of the range can be very different than the middle or bottom. For example, National Taiwan University^[<https://www.ntu.edu.tw/english/>] is the most prestigious university in Taiwan, but Chang Gung University^[<https://www.cgu.edu.tw/?Lang=en>] is also excellent in medical education. However, many people in the US have heard of the former but not the latter. This does not mean you should always choose the most prestigious college; many other factors should also be considered, such as tuition and majors offered at that college.  

We consider the following subcategories in the entrance exam scores:     

* **HighSchool_PR** ranges: 80-89, 90-94, 95-99
* **College_Score** ranges: 60-64, 65-69, 70-75  

Let's investigate the univariate distributions of **HighSchool_PR** and **College_Score** respectively, then we move on to explore the bivariate relationship between the two sets of scores. It is good practice to start with one variable at a time, even when we are mainly interested in the relationship between the two variables.  

## High School Entrance Exam Scores (Percentile Rank) at least 80 {#HighSchool-PR-80-up}

We use the `R` function `table` to show the frequency of each **HighSchool_PR** value that is at least 80, and we have 163 values in total. In the table below, the first row is the PR (percentile rank), and the second row is the counts. Although we truncated the **HighSchool_PR** to 80 and above, the distribution is still left-skewed. The **HighSchool_PR** 99 has the highest counts, followed by **HighSchool_PR** 97 and **HighSchool_PR** 98.

```{r high-school-pr-sequential}
HS_PR_seg = data_corr$HighSchool_PR[which(data_corr$HS_80up == TRUE)]
length(HS_PR_seg)
```

```{r high-school-pr-table}
table(HS_PR_seg)
```

Or if you prefer a histogram, we can also create one.

```{r high-school-pr-seg-hist}
hist(HS_PR_seg, xlab="HighSchool_PR",
     main="Histogram of HighSchool_PR 80 and above")
```

Therefore, we create the breakdown of the **HighSchool_PR** ranges: 80-89, 90-94, 95-99. There are 49 records in the 80-89 range, 44 records in 90-94, and 70 records in 95-99. We divided the 90-99 range into 90-94 and 95-99, but the number of **HighSchool_PR** records in the 95-99 range is still higher than any of the other two categories.  

However, it is not a good idea to further divide the 95-99 range into 95-97 and 98-99, due to the lack of geographic information. In Taipei, the high school enrollment is extremely competitive. Students with **HighSchool_PR** 95 and those with **HighSchool_PR** 99 would get admitted to high schools of different rankings.^[https://w199584.pixnet.net/blog/post/28321887] But in other parts of Taiwan, most students with **HighSchool_PR** at least 95 would already qualify for the top local high school, and some rural parts even require a lower **HighSchool_PR** to get into the county's top high school.

```{r high-school-pr-segments}
HS80to89 = length(which(HS_PR_seg >= 80 & HS_PR_seg <= 89))
HS90to94 = length(which(HS_PR_seg >= 90 & HS_PR_seg <= 94))
HS95to99 = length(which(HS_PR_seg >= 95 & HS_PR_seg <= 99))
```

```{r high-school-pr-segments-print}
print(paste("HighSchool_PR 80-89:",HS80to89))
print(paste("HighSchool_PR 90-94:",HS90to94))
print(paste("HighSchool_PR 95-99:",HS95to99))
```

## College Entrance Exam Scores at least 60 {#college-60-up}

Similar to the previous section, we also show the frequency of each **College_Score** value that is at least 60. The total counts is 128, fewer than the 163 counts with **HighSchool_PR** 80 or above. The distribution is relatively uniform for **College_Score** values between 60 and 73, with a steep decline in the counts of **College_Score** 74 and 75 (max possible score). On the college entrance exam, only four respondents scored 74 and two scored 75. According to the historical statistics of the college entrance exam in Taiwan, **College_Score** 74 and 75 account for approximately 0.2% of all test takers each year, which is quite a small percentage.    

```{r college-score-sequential}
CS_Score_seg = data_corr$College_Score[which(data_corr$CS_60up == TRUE)]
length(CS_Score_seg)
```

```{r college-score-table}
table(CS_Score_seg)
```

Before we display the histogram, let's create a table to (approximately) convert **College_Score** into PR (Percentile Rank) using 2001-2014 data.^[https://web.archive.org/web/20150207042900/http://www.ceec.edu.tw/AbilityExam/AbilityExamStat.htm] This gives readers an understanding of what percentage of test takers (high school students in grade 12) get what scores. For example, a **College_Score** of 70 would be at the 98.5th percentile, i.e., PR 98.5.

```{r college-score-pr}
college_score = c(60,65,70,74,75)
college_pr = c(88, 95, 98.5, 99.9, 100)

data.frame(college_score, college_pr)
```

Here is the histogram of the **College_Score** values 60 and above.

```{r college-score-seg-hist}
hist(CS_Score_seg, xlab="College_Score (max possible is 75)",
     main="Histogram of College_Score 60 and above")
```

We also create the breakdown of the **College_Score** ranges: 60-64, 65-69, 70-75. There are 36 records in the 60-64 range, 47 records in 65-69, and 45 records in 70-75. This is also relatively more uniform than the **HighSchool_PR** breakdown (49, 44, and 70 records each).

```{r college-score-segments}
CS60to64 = length(which(CS_Score_seg >= 60 & CS_Score_seg <= 64))
CS65to69 = length(which(CS_Score_seg >= 65 & CS_Score_seg <= 69))
CS70to75 = length(which(CS_Score_seg >= 70 & CS_Score_seg <= 75))
```

```{r college-score-segments-print}
print(paste("College_Score 60-64:",CS60to64))
print(paste("College_Score 65-69:",CS65to69))
print(paste("College_Score 70-75:",CS70to75))
```

