---
title: "PTT Analysis of Entrance Exam Scores in Taiwan"
subtitle: "A Statistical Approach with R Code"
author:
  - Christine P. Chai
  - Microsoft
  - cpchai21@gmail.com
date: \today
output: 
        pdf_document:
                number_sections: true
                citation_package: natbib
header-includes: 
  - \renewcommand{\and}{\\}
bibliography: references.bib
biblio-style: apalike
link-citations: yes
---

\renewcommand{\cite}{\citep}

```{r latex-cite-command, include=FALSE}
# %\let\cite\citep
# % from \citep to \cite to cite in author style, e.g. [Mule, 2008]

# % \bibliographystyle{plainnat}
# %\citep: citation in parentheses, e.g. [Mule, 2008]
# %\citet: citation as author, e.g. Mule [2008]
# %\cite: citation as author, \citet by default 
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

def.chunk.hook  <- knitr::knit_hooks$get("chunk")
knitr::knit_hooks$set(chunk = function(x, options) {
  x <- def.chunk.hook(x, options)
  ifelse(options$size != "normalsize", paste0("\n \\", options$size,"\n\n", x, "\n\n \\normalsize"), x)
})
```

Ongoing work since 2019.

# Executive Summary {.unnumbered}

\textcolor{red}{Write something here}

# Disclaimer {.unnumbered}

The opinions and views expressed in this manuscript are those of the author and do not necessarily state or reflect those of Microsoft. The author has a statistics and engineering background, without any training in middle school or high school teaching.

# Introduction {#intro}

```{r include-intro, child = '01-Introduction.Rmd'}
```

# Background {#background}

\section*{\textcolor{red}{Unfinished below}}

Elaborate on the following to enhance the background context:

- College entrance exam GSAT and AST: New rules in 2022 (not related to COVID)

- COVID-related rules for students who could not attend the exam  
Start writing about 2021 for both exams, because 2022 is still ongoing  

- If we continue writing this manuscript in 2023, we absolutely need to add the recent changes in 2021 and 2022.  

\textcolor{red}{This is the file we want to move content into.}  

\section*{\textcolor{red}{Unfinished above}}

```{r include-background, child = '02-Background.Rmd'}
```

# Data Description {#data}

\section*{\textcolor{red}{Unfinished below}}

\textcolor{red}{Definitely need to fix both data sections}   

\textcolor{red}{This is the file we want to move content into.}  

\section*{\textcolor{red}{Unfinished below}}  

```{r include-data-description, child = '03-Data-Description.Rmd'}
```

# Exploratory Data Analysis {#eda}

```{r include-exploratory, child = '04-Exploratory.Rmd'}
```

# Linear Regression {#linear-reg}

```{r include-linear-reg, child = '05-Linear-Regression.Rmd'}
```

# Top Scorers: A Closer Look {#explore-top}

```{r include-linear-reg, child = '06-Top-Scorers-01-Univariate.Rmd'}
```

## Bivariate Exploration of Top Scorers {#bivariate-top-scorers}

```{r include-chisq-test, child = '06-Top-Scorers-02-Bivariate.Rmd'}
```

# Logistic Regression {#logit-reg}

```{r include-logit-reg, child = '07-Logistic-Regression.Rmd'}
```

# Model Validation: In-Sample Prediction {#validation}

```{r include-in-sample-00, child = '08-In-Sample-00-Foreword.Rmd'}
```

## Implementation of In-Sample Prediction {#in-sample}

```{r include-in-sample-01, child = '08-In-Sample-01-Implementation.Rmd'}
```

## Interpretation of Confusion Matrix {#interpretation}

\section*{\textcolor{red}{Unfinished below}}  

What about ROC (receiver operating characteristic) and AUC (area under the curve) with various probability thresholds?  

In many cases, it is not obvious to determine the probability threshold to predict a datapoint to be positive.  

But ROC requires the definitions of TPR and FPR first.  

You don't have to do the full implementation, just write a few lines as a pointer to the concepts.  

\textcolor{red}{This is the file we want to move content into.}  

\section*{\textcolor{red}{Unfinished above}}  

```{r include-in-sample-02, child = '08-In-Sample-02-Confusion-Matrix.Rmd'}
```

## Breakdown by High School Entrance Exam Scores {#another-breakdown}

\section*{\textcolor{red}{Unfinished below}}  

Add my thoughts about categorical data analysis (e.g. proportional hazards model?!)  

Example: When **HighSchool_PR** is increased from Category X to Category Y, how does it affect the categories of **College_Score**?  

You don't have to do the full implementation, just write a few lines as a pointer to the concepts.  

\textcolor{red}{This is the file we want to move content into.}  

\section*{\textcolor{red}{Unfinished above}} 

```{r include-in-sample-03, child = '08-In-Sample-03-Score-Breakdown.Rmd'}
```

# Model Validation: Out-of-Sample Prediction {#out-of-sample}

```{r include-out-sample-00, child = '09-Out-Sample-00-Foreword.Rmd'}
```

## Separate Training and Testing Datasets

\section*{\textcolor{red}{Unfinished below}}  

Add a remark on hyperparameter tuning for random forests.  

Hyperparameter tuning is also necessary for ridge regression and Lasso regression, both of which contain a penalty term to regulate the number of coefficients in the model.  

\section*{\textcolor{red}{Unfinished above}}  

```{r include-out-sample-01, child = '09-Out-Sample-01-Sep-Explain.Rmd'}
```

### Implementation {#train-test-demo}

```{r include-out-sample-02, child = '09-Out-Sample-02-Sep-Implement.Rmd'}
```

### Organizing the Code for Reusability

```{r include-out-sample-03, child = '09-Out-Sample-03-Sep-Organize.Rmd'}
```

## Cross Validation

```{r include-out-sample-04, child = '09-Out-Sample-04-Cross-Validation.Rmd'}
```

### K-fold Cross Validation

```{r include-out-sample-05, child = '09-Out-Sample-05-K-Fold.Rmd'}
```

### Leave-one-out Cross Validation

```{r include-out-sample-06, child = '09-Out-Sample-06-Leave-One-Out.Rmd'}
```

## Comparison of Results {#cmp-results}

```{r include-out-sample-07, child = '09-Out-Sample-07-Comparison-Results.Rmd'}
```

# Recap of the Project {#recap}

```{r include-project-recap, child = '10-Project-Recap.Rmd'}
```

# Recommended Resources for Learning {#resources}

```{r include-additional-resources, child = '11-Additional-Resources.Rmd'}
```

# Final: Personal Scores and Remarks {#personal-remarks}

```{r include-personal, child = '12-Personal-Scores-Remarks.Rmd'}
```

# Acknowledgments {.unnumbered}

The author would like to thank Dr. Mine Cetinkaya-Rundel and Dr. David Banks at Duke University; they both motivated the author to teach statistics and create reproducible work in $\mathsf{R}$. The author is also grateful to her Microsoft colleagues Smit Patel and Dylan Stout for troubleshooting GitHub issues.  

The author would also like to acknowledge Dr. Cliburn Chan and Dr. Janice McCarthy for introducing her to GitHub in the statistical computation course at Duke University. This provided her the foundations to use GitHub as a modern version control system in the first place.  

Finally, the author gives a special mention to her significant other, Hugh Hendrickson, for all his support in the author's professional career development.



# References {.unnumbered #references}