---
title: "PTT Analysis of Entrance Exam Scores in Taiwan"
subtitle: "A Statistical Approach with R Code"
author:
  - Christine P. Chai
  - Microsoft
  - cpchai21@gmail.com
date: \today
output: 
        pdf_document:
                number_sections: true
                citation_package: natbib
header-includes: 
  - \renewcommand{\and}{\\}
bibliography: references.bib
biblio-style: apalike
link-citations: yes
---

\renewcommand{\cite}{\citep}

```{r latex-cite-command, include=FALSE}
# %\let\cite\citep
# % from \citep to \cite to cite in author style, e.g. [Mule, 2008]

# % \bibliographystyle{plainnat}
# %\citep: citation in parentheses, e.g. [Mule, 2008]
# %\citet: citation as author, e.g. Mule [2008]
# %\cite: citation as author, \citet by default 
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Ongoing work since 2019.

# Executive Summary {.unnumbered}

\textcolor{red}{Write something here}

# Disclaimer {.unnumbered}

The opinions and views expressed in this manuscript are those of the author and do not necessarily state or reflect those of Microsoft. The author has a statistics and engineering background, without any training in middle school or high school teaching.

# Introduction {#intro}

\section*{\textcolor{red}{Unfinished below}}

Consider making the introduction statistics-oriented, because we would like to focus on the statistical methodology. Most important is how to handle a real-time application problem.  

State this project as open-source and provide the GitHub link somewhere in the manuscript. Note that the name of GitHub repo may be changed later.  

Edit the paragraph of this manuscript's structure! Use cross-references for each chapter.  

\section*{\textcolor{red}{Unfinished above}}

```{r include-intro, child = '01-Introduction.Rmd'}
```

# Background {#background}

\section*{\textcolor{red}{Unfinished below}}

Elaborate on the following to enhance the background context:

- Starting in 2019, high school students are required to maintain a portfolio each semester to document their learning outcomes.^[<https://www.108epo.com/courses.php>]

- College entrance exam GSAT and AST: New rules in 2022 (not related to COVID)

- COVID-related rules for students who could not attend the exam  
Start writing about 2021 for both exams, because 2022 is still ongoing

\section*{\textcolor{red}{Unfinished above}}

```{r include-background, child = '02-Background.Rmd'}
```

# Data Description {#data}

\section*{\textcolor{red}{Unfinished below}}

Add the new PTT image and state **when** the registration restrictions were removed.  

Note that from 2018 to 2021, PTT limited new account registrations to only people with an email address from National Taiwan University.^[Screenshot obtained on May 26, 2020: <https://imgur.com/33fwrGH>]   

The good news is that this restriction has been removed.^[Screenshot obtained on February 5, 2022: <https://imgur.com/kaTsRpn>] As of 2022, anyone can register for an account on PTT with email^[<https://www.techbang.com/posts/88695-ptt-registration>] or mobile phone authentication^[<https://free.com.tw/ptt-id-registration-aotp/>].  

Also, list several mobile apps for readers to access PTT on a smart phone. Common apps include BePTT, Mo PTT, JPTT, and PiTT.

\section*{\textcolor{red}{Unfinished above}}

```{r include-data-description, child = '03-Data-Description.Rmd'}
```

# Exploratory Data Analysis {#eda}

```{r include-exploratory, child = '04-Exploratory.Rmd'}
```

# Linear Regression {#linear-reg}

```{r include-linear-reg, child = '05-Linear-Regression.Rmd'}
```

# Top Scorers: A Closer Look {#explore-top}

\section*{\textcolor{red}{Unfinished below}}  

Bivariate exploration subsection:  

Add my thoughts about categorical data analysis (e.g. proportional hazards model?!)  

\section*{\textcolor{red}{Unfinished above}}  

```{r include-linear-reg, child = '06-Top-Scorers.Rmd'}
```

# Logistic Regression {#logit-reg}

```{r include-logit-reg, child = '07-Logistic-Regression.Rmd'}
```

# Model Validation: In-Sample Prediction {#validation}

```{r include-in-sample-00, child = '08-In-Sample-00-Foreword.Rmd'}
```

## Implementation of In-Sample Prediction {#in-sample}

```{r include-in-sample-01, child = '08-In-Sample-01-Implementation.Rmd'}
```

## Interpretation of Confusion Matrix {#interpretation}

```{r include-in-sample-02, child = '08-In-Sample-02-Confusion-Matrix.Rmd'}
```

## Breakdown by High School Entrance Exam Scores {#another-breakdown}

```{r include-in-sample-03, child = '08-In-Sample-03-Score-Breakdown.Rmd'}
```

# Model Validation: Out-of-Sample Prediction {#out-of-sample}

```{r include-out-sample-00, child = '09-Out-Sample-00-Foreword.Rmd'}
```

## Separate Training and Testing Datasets

```{r include-out-sample-01, child = '09-Out-Sample-01-Sep-Explain.Rmd'}
```

### Implementation {#train-test-demo}

```{r include-out-sample-02, child = '09-Out-Sample-02-Sep-Implement.Rmd'}
```

### Organizing the Code for Reusability

```{r include-out-sample-03, child = '09-Out-Sample-03-Sep-Organize.Rmd'}
```

## Cross Validation

```{r include-out-sample-04, child = '09-Out-Sample-04-Cross-Validation.Rmd'}
```

### K-fold Cross Validation

```{r include-out-sample-05, child = '09-Out-Sample-05-K-Fold.Rmd'}
```

### Leave-one-out Cross Validation

```{r include-out-sample-06, child = '09-Out-Sample-06-Leave-One-Out.Rmd'}
```

## Comparison of Results {#cmp-results}

```{r include-out-sample-07, child = '09-Out-Sample-07-Comparison-Results.Rmd'}
```

# Recap of the Project {#recap}

```{r include-project-recap, child = '10-Project-Recap.Rmd'}
```

# Recommended Resources for Learning {#resources}

```{r include-additional-resources, child = '11-Additional-Resources.Rmd'}
```

# Final: Personal Scores and Remarks {#personal-remarks}

```{r include-personal, child = '12-Personal-Scores-Remarks.Rmd'}
```

# Acknowledgments {.unnumbered}

The author would like to thank Dr. Mine Cetinkaya-Rundel and Dr. David Banks at Duke University; they both motivated the author to teach statistics and create reproducible work in $\mathsf{R}$. The author is also grateful to her Microsoft colleagues Smit Patel and Dylan Stout for troubleshooting GitHub issues.  

The author would also like to acknowledge Dr. Cliburn Chan and Dr. Janice McCarthy for introducing her to GitHub in the statistical computation course at Duke University. This provided her the foundations to use GitHub as a modern version control system in the first place.  

Finally, the author gives a special mention to her significant other, Hugh Hendrickson, for all his support in the author's professional career development.

# Appendix {.unnumbered}

\section*{\textcolor{red}{Unfinished below}}

\textcolor{red}{Consider moving the Appendix into Additional Resources}  

```{r include-appendix, child = 'Appendix.Rmd'}
```

This document was created with $\mathsf{R}$ Markdown \cite{r-markdown-pkg} in RStudio.^[<https://rmarkdown.rstudio.com/articles_intro.html>] $\mathsf{R}$ Markdown has a framework to directly incorporate code and graphs in the report. This not only reduces the errors from copy-pasting tables and plots, but also ensures reproducibility of the data analysis \cite{baumer2014r-markdown}. $\mathsf{R}$ Markdown can generate reports in PDF, HTML, or Microsoft Word formats. $\mathsf{R}$ Markdown also provides a template for presentation slides, and the output can be in HTML, PDF, or Microsoft PowerPoint files.    

Also consider: `bookdown` \cite{r-bookdown}  

The graphs in this manuscript are mainly for demonstration; they are created using the basic `plot` function as the default in RStudio. For advanced data visualization, we highly encourage readers to use the $\mathsf{R}$ package `ggplot2` \cite{ggplot2}. This package incorporates data visualization theory and provides much more capability to convey the message in a graph.       

Consider adding books about data visualization theory, citations needed  

- *The Grammar of Graphics* \cite{wilkinson2012grammar}: mainly for `ggplot2`

- *Semiology of Graphics: Diagrams, Networks, Maps* \cite{bertin1983semiology}: deep dive of data visualization theory

- *Visualize This* \cite{yau2011visualize}: simple, hands-on guide for data visualization  

For advanced operations in $\mathsf{R}$, we recommend the $\mathsf{R}$ package `tidyverse` \cite{r-tidyverse} because it provides a fast, efficient, and organized workflow for general data modeling.^[<https://www.r-bloggers.com/2018/09/why-learn-the-tidyverse/>] The book \textit{$\mathsf{R}$ for Data Science: Import, Tidy, Transform, Visualize, and Model Data} \cite{wickham2016r} demonstrates an end-to-end cycle of using `tidyverse` in data science.  

The code is also for demonstration. The author makes every effort to ensure correctness and readability of the code, but errors and inconsistencies are inevitable. Since this is an open-source project on GitHub,^[<https://github.com/star1327p/Exam-Scores-PTT>], readers are welcome to make a pull request and/or report any issues. To learn more about better coding styles, we recommend the computer science books *Clean Code* \cite{martin2009clean} and *Code Complete* \cite{mcconnell2004code}.  

Talk about Python, `numpy`, and `scikit-learn`, etc?!

# References {.unnumbered #references}