---
title: "Statistical Analysis of High School and College Entrance Exam Scores in Taiwan with Online Data"
subtitle: "A Fully Reproducible Approach with R Markdown"
author:
  - Christine P. Chai
  - cpchai21@gmail.com
date: \today
output: 
        pdf_document:
                number_sections: true
                citation_package: natbib
header-includes: 
  - \renewcommand{\and}{\\}
bibliography: references.bib
biblio-style: apalike
link-citations: yes
---

\renewcommand{\cite}{\citep}

```{r latex-cite-command, include=FALSE}
# %\let\cite\citep
# % from \citep to \cite to cite in author style, e.g. [Mule, 2008]

# % \bibliographystyle{plainnat}
# %\citep: citation in parentheses, e.g. [Mule, 2008]
# %\citet: citation as author, e.g. Mule [2008]
# %\cite: citation as author, \citet by default 
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

def.chunk.hook  <- knitr::knit_hooks$get("chunk")
knitr::knit_hooks$set(chunk = function(x, options) {
  x <- def.chunk.hook(x, options)
  ifelse(options$size != "normalsize", paste0("\n \\", options$size,"\n\n", x, "\n\n \\normalsize"), x)
})
```

\section*{\textcolor{red}{Unfinished below}}

Ongoing work since 2019.   

To-Do:

1. Need to add more references, instead of just URLs. Especially on Taiwan's education system, which I have very few items listed in the References. Most of my references are about statistics and/or data science itself.

- Tentative goal: Add enough references for the PDF to have at least 90 pages.

- The Introduction section needs to be stronger and more convincing.

- Definitely need more references for the Background section.


Done:

- Proofread the content myself (85+ pages).

- Due to the discontinuation of **xuite.net** effective August 31, 2023,^[<https://mrmad.com.tw/xuite-close>] I need to replace these footnote links with something else. 

- Checked the validity of each URL.

- For any URL that is long and/or contains a hash string, shorten the URL to be concise.

- Avoid using gender-indicative pronouns like "he/she", "him/her", or "his/her". Change it to "they", "their", or something not related to gender.

- Section \ref{College_GSAT}: Add the changes in 2022 -- GSAT contains mathematics A and mathematics B.^[<https://bit.ly/4bEoQ0y>]  


Continue:

2. Also check the references in detail. 

3. Ask some friends (preferably current high school teachers) to verify the correctness of the high school and college entrance rules in Taiwan.  

4. Ask other friends for feedback on my report.

5. Revise the report according to the feedback I receive.

6. Recheck everything one more time.

7. Obtain approval to list my new affiliation with the disclaimer.

8. Finally, publish this article to ResearchGate.

\section*{\textcolor{red}{Unfinished above}}

# Executive Summary {.unnumbered}

In this project, we investigate the relationship between high school and college entrance exam scores in Taiwan using self-reported data online. The goal is to demonstrate reproducible statistical analysis in $\mathsf{R}$ Markdown and LaTeX to PDF. We start by exploring the data to formulate the problem statement and identify the appropriate model, which can be an iterative process. We eventually decided on a binary outcome of the college entrance exams, and hence implemented a logistic regression to fit the model. We also validated the model with out-of-sample prediction methods, including cross validation as well as separate training and testing sets. Finally, we used some metrics to evaluate the model performance. The end-to-end data analysis can be compiled with a single button in RStudio, without copying-and-pasting output from one program to another.   


# Disclaimer {.unnumbered}

The opinions and views expressed in this manuscript are those of the author, and do not necessarily state or reflect those of any institution or government entity. The author has a statistics and engineering background, without any training in middle school or high school teaching.

# Introduction {#intro}

```{r include-intro, child = '01-Introduction.Rmd'}
```

# Background {#background}

```{r include-background, child = '02-Background.Rmd'}
```

# Data Description {#data}

```{r include-data-description, child = '03-Data-Description.Rmd'}
```

# Exploratory Data Analysis {#eda}

```{r include-exploratory, child = '04-Exploratory.Rmd'}
```

# Linear Regression {#linear-reg}

```{r include-linear-reg, child = '05-Linear-Regression.Rmd'}
```

# Top Scorers: A Closer Look {#explore-top}

```{r include-linear-reg, child = '06-Top-Scorers-01-Univariate.Rmd'}
```

## Bivariate Exploration of Top Scorers {#bivariate-top-scorers}

```{r include-chisq-test, child = '06-Top-Scorers-02-Bivariate.Rmd'}
```

# Logistic Regression {#logit-reg}

```{r include-logit-reg, child = '07-Logistic-Regression.Rmd'}
```

# Model Validation: In-Sample Prediction {#validation}

```{r include-in-sample-00, child = '08-In-Sample-00-Foreword.Rmd'}
```

## Implementation of In-Sample Prediction {#in-sample}

```{r include-in-sample-01, child = '08-In-Sample-01-Implementation.Rmd'}
```

## Interpretation of Confusion Matrix {#interpretation}

```{r include-in-sample-02, child = '08-In-Sample-02-Confusion-Matrix.Rmd'}
```

## Breakdown by High School Entrance Exam Scores {#another-breakdown}

```{r include-in-sample-03, child = '08-In-Sample-03-Score-Breakdown.Rmd'}
```

# Model Validation: Out-of-Sample Prediction {#out-of-sample}

```{r include-out-sample-00, child = '09-Out-Sample-00-Foreword.Rmd'}
```

## Separate Training and Testing Datasets {#sep-train-test}

```{r include-out-sample-01, child = '09-Out-Sample-01-Sep-Explain.Rmd'}
```

### Implementation {#train-test-demo}

```{r include-out-sample-02, child = '09-Out-Sample-02-Sep-Implement.Rmd'}
```

### Organizing the Code for Reusability {#org-code-reuse}

```{r include-out-sample-03, child = '09-Out-Sample-03-Sep-Organize.Rmd'}
```

## Cross Validation {#cross-validation}

```{r include-out-sample-04, child = '09-Out-Sample-04-Cross-Validation.Rmd'}
```

### K-fold Cross Validation {#k-fold}

```{r include-out-sample-05, child = '09-Out-Sample-05-K-Fold.Rmd'}
```

### Leave-one-out Cross Validation {#leave-one-out}

```{r include-out-sample-06, child = '09-Out-Sample-06-Leave-One-Out.Rmd'}
```

## Comparison of Results {#cmp-results}

```{r include-out-sample-07, child = '09-Out-Sample-07-Comparison-Results.Rmd'}
```

# Model Metrics: ROC and AUC {#roc-auc}

```{r include-roc-auc-00, child = '10-Model-Metrics-00-Foreword.Rmd'}
```

## Demonstrative ROC Curve {#roc-demo}

```{r include-roc-auc-01, child = '10-Model-Metrics-01-Demo.Rmd'}
```

## Data Observation and Processing {#roc-prep} 

```{r include-roc-auc-02, child = '10-Model-Metrics-02-Observation.Rmd'}
```

## Implementation and Results {#roc-auc-results}

```{r include-roc-auc-03, child = '10-Model-Metrics-03-ROC-AUC.Rmd'}
```

# Recap of the Project {#recap}

```{r include-project-recap, child = '11-Project-Recap.Rmd'}
```

# Recommended Resources for Learning {#resources}

```{r include-additional-resources, child = '12-Additional-Resources.Rmd'}
```

# Personal Scores and Remarks {#personal-remarks}

```{r include-personal, child = '13-Personal-Scores-Remarks.Rmd'}
```

# Final Words {.unnumbered}

```{r include-final-words, child = '14-Final-Words.Rmd'}
```

# Acknowledgments {.unnumbered}

The author would like to thank Dr. Mine Cetinkaya-Rundel and Dr. David Banks at Duke University; they both motivated the author to teach statistics and create reproducible work in $\mathsf{R}$. The author is also grateful to her former Microsoft colleagues Smit Patel and Dylan Stout for troubleshooting GitHub issues.  

The author would also like to acknowledge Dr. Cliburn Chan and Dr. Janice McCarthy for introducing her to GitHub in the statistical computation course at Duke University.^[<https://people.duke.edu/~ccc14/sta-663-2015/>] This provided her the foundations to use GitHub as a modern version control system in the first place.  

Finally, the author gives a special mention to her significant other, Hugh Hendrickson, for all his support in the author's professional career development.


# References {.unnumbered #references}