---
title: "PTT Analysis of Entrance Exam Scores in Taiwan"
subtitle: "A Statistical Approach with R Code"
author:
  - Christine P. Chai
  - Microsoft
  - cpchai21@gmail.com
date: \today
output: 
        pdf_document:
                number_sections: true
                citation_package: natbib
header-includes: 
  - \renewcommand{\and}{\\}
bibliography: references.bib
biblio-style: apalike
link-citations: yes
---

\renewcommand{\cite}{\citep}

```{r latex-cite-command, include=FALSE}
# %\let\cite\citep
# % from \citep to \cite to cite in author style, e.g. [Mule, 2008]

# % \bibliographystyle{plainnat}
# %\citep: citation in parentheses, e.g. [Mule, 2008]
# %\citet: citation as author, e.g. Mule [2008]
# %\cite: citation as author, \citet by default 
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Ongoing work since 2019.

# Executive Summary {.unnumbered}

\textcolor{red}{Write something here}

# Disclaimer {.unnumbered}

The opinions and views expressed in this manuscript are those of the author and do not necessarily state or reflect those of Microsoft. The author has a statistics and engineering background, without any training in middle school or high school teaching.

# Introduction {#intro}

\section*{\textcolor{red}{Unfinished below}}

Consider making the introduction statistics-oriented, because we would like to focus on the statistical methodology. Most important is how to handle a real-time application problem.  

State this project as open-source and provide the GitHub link somewhere in the manuscript. Note that the name of GitHub repo may be changed later.  

Edit the paragraph of this manuscript's structure! Use cross-references for each chapter.  

\section*{\textcolor{red}{Unfinished above}}

```{r include-intro, child = '01-Introduction.Rmd'}
```

# Background {#background}

\section*{\textcolor{red}{Unfinished below}}

Elaborate on the following to enhance the background context:

- Starting in 2019, high school students are required to maintain a portfolio each semester to document their learning outcomes.^[<https://www.108epo.com/courses.php>]

- College entrance exam GSAT and AST: New rules in 2022 (not related to COVID)

- COVID-related rules for students who could not attend the exam  
Start writing about 2021 for both exams, because 2022 is still ongoing

\section*{\textcolor{red}{Unfinished above}}

```{r include-background, child = '02-Background.Rmd'}
```

# Data Description {#data}

\section*{\textcolor{red}{Unfinished below}}

Add the new PTT image and state **when** the registration restrictions were removed.  

Note that from 2018 to 2021, PTT limited new account registrations to only people with an email address from National Taiwan University.^[Screenshot obtained on May 26, 2020: <https://imgur.com/33fwrGH>]   

The good news is that this restriction has been removed.^[Screenshot obtained on February 5, 2022: <https://imgur.com/kaTsRpn>] As of 2022, anyone can register for an account on PTT with email^[<https://www.techbang.com/posts/88695-ptt-registration>] or mobile phone authentication^[<https://free.com.tw/ptt-id-registration-aotp/>].  

Also, list several mobile apps for readers to access PTT on a smart phone. Common apps include BePTT, Mo PTT, JPTT, and PiTT.

\section*{\textcolor{red}{Unfinished above}}

```{r include-data-description, child = '03-Data-Description.Rmd'}
```

# Exploratory Data Analysis {#eda}

```{r include-exploratory, child = '04-Exploratory.Rmd'}
```

# Linear Regression {#linear-reg}

```{r include-linear-reg, child = '05-Linear-Regression.Rmd'}
```

# Top Scorers: A Closer Look {#explore-top}

\section*{\textcolor{red}{Unfinished below}}  

Bivariate exploration subsection:  

Add my thoughts about categorical data analysis (e.g. proportional hazards model?!)  

\section*{\textcolor{red}{Unfinished above}}  

```{r include-linear-reg, child = '06-Top-Scorers.Rmd'}
```

# Logistic Regression {#logit-reg}

```{r include-logit-reg, child = '07-Logistic-Regression.Rmd'}
```

# Model Validation: In-Sample Prediction {#validation}

```{r include-in-sample-00, child = '08-In-Sample-00-Foreword.Rmd'}
```

## Implementation of In-Sample Prediction {#in-sample}

```{r include-in-sample-01, child = '08-In-Sample-01-Implementation.Rmd'}
```

## Interpretation of Confusion Matrix {#interpretation}

```{r include-in-sample-02, child = '08-In-Sample-02-Confusion-Matrix.Rmd'}
```

## Breakdown by High School Entrance Exam Scores {#another-breakdown}

```{r include-in-sample-03, child = '08-In-Sample-03-Score-Breakdown.Rmd'}
```

# Model Validation: Out-of-Sample Prediction {#out-of-sample}

```{r include-out-sample-00, child = '09-Out-Sample-00-Foreword.Rmd'}
```

## Separate Training and Testing Datasets

```{r include-out-sample-01, child = '09-Out-Sample-01-Sep-Explain.Rmd'}
```

### Implementation {#train-test-demo}

```{r include-out-sample-02, child = '09-Out-Sample-02-Sep-Implement.Rmd'}
```

### Organizing the Code for Reusability

```{r include-out-sample-03, child = '09-Out-Sample-03-Sep-Organize.Rmd'}
```

## Cross Validation

```{r include-out-sample-04, child = '09-Out-Sample-04-Cross-Validation.Rmd'}
```

### K-fold Cross Validation

```{r include-out-sample-05, child = '09-Out-Sample-05-K-Fold.Rmd'}
```

### Leave-one-out Cross Validation

```{r include-out-sample-06, child = '09-Out-Sample-06-Leave-One-Out.Rmd'}
```

## Comparison of Results {#cmp-results}

```{r include-out-sample-07, child = '09-Out-Sample-07-Comparison-Results.Rmd'}
```

# Recap of the Project {#recap}

```{r include-project-recap, child = '10-Project-Recap.Rmd'}
```

# Recommended Resources for Learning {#resources}

\section*{\textcolor{red}{Unfinished below}}

Add the foreword and the following subsections:

- Resources for Learning Statistics

- Resources for Data Science and Programming  

```{r include-additional-resources, child = '11-Additional-Resources.Rmd'}
```

State that statistics is the core of data analysis?!  

## Resources for Learning Statistics

\textcolor{red}{Need to elaborate on the first paragraph. Cannot just write one line.}  

Here are some resources we recommend for the path of learning statistics:  
  
The Statistics 101 course covers the fundamental statistical concepts for students to perform data analysis, and the programming component is often included. For readers who would like a refresher, we recommend the textbook *OpenIntro Statistics* \cite{diez2019openintro} and the online course series *Statistics with `R` Specialization* on Coursera.^[<https://www.coursera.org/specializations/statistics>] Basic concepts like hypothesis testing and confidence intervals are also asked in technical interviews for data science jobs. Beware! For an overview of Bayesian Statistics, we recommend the *Bayesian Statistics* online course on Coursera,^[<https://www.coursera.org/learn/bayesian>] along with the open-source textbook *An Introduction to Bayesian Thinking* \cite{clyde2018bayesian101}. At this stage, knowledge of Calculus is helpful but not required.  
  
As the next step beyond the introductory level, we suggest reading *The Statistical Sleuth: A Course in Methods of Data Analysis* \cite{ramsey2013statistical}, which is the textbook for undergraduate-level regression analysis at Duke Statistical Science.^[<https://www2.stat.duke.edu/courses/Fall18/sta210.001/>] The book covers intermediate topics such as ANOVA (Analysis of Variance) and multiple linear regression. It also provides data files for case studies and exercises.^[<http://www.statisticalsleuth.com/>]   


\textcolor{red}{Elaborate on each textbook and state the math prerequisites}  

Also add textbooks from Duke Stat courses: STA230, STA250, STA611  

- *Statistics: Unlocking the Power of Data* \cite{lock2020statistics}: More applied

- *Probability and Statistics* \cite{degroot2012probability}: More theoretical, and Calculus is a prerequisite.  

When the readers go further in statistics, they may realize that college-level mathematics becomes increasingly important. Formulas and equations will gradually appear in the curriculum, especially Calculus for continuous functions and matrix algebra for discrete datapoints.   

For the advanced readers, we recommend the following graduate level statistics textbooks. Note that multivariate Calculus and matrix algebra are absolutely necessary.      

- *A First Course in Bayesian Statistical Methods* \cite{hoff2009first}: Bayesian statistics  

- *A Modern Approach to Regression with `R`* \cite{sheather2009modern}: Linear models

- *Statistical Inference* \cite{casella2021statistical}: The theory behind statistical concepts such as maximum likelihood estimation

- *Categorical Data Analysis* \cite{agresti2003categorical}: Logistic regression and others?!

There are many other high-quality statistics textbooks, and we selected these as a starting point. If the readers are interested in a data science career, learning statistics builds the foundations and expands your toolbox to generate insights from data.^[<https://towardsdatascience.com/the-difference-between-data-science-and-statistics-168c7062c201>] We also recommend exploring Kaggle datasets^[<https://www.kaggle.com/datasets>] for hands-on practice. To get closer to analyzing real-world data, we encourage the readers to attend data hackathons, contribute to open-source projects, and/or volunteer technical skills at nonprofit organizations like Statistics Without Borders.^[<https://www.statisticswithoutborders.org/>] These activities are not a comprehensive list, and please feel free to get creative and carve your own professional path forward.    

## Resources for Data Science and Programming

Since we discussed plenty of statistics resources in the previous section, we would like to provide a few pointers to **data science**, one of the hottest fields in tech. Programming is also essential to handle data, so we introduce some practical tools that readers can incorporate them into their data science workflow. (Don't get overwhelmed with the massive selection of tools; you can try one at a time.)  

This document is created with $\mathsf{R}$ Markdown \cite{r-markdown-pkg},^[<https://rmarkdown.rstudio.com/articles_intro.html>] which has a framework to directly incorporate code and graphs in the report. This not only reduces the errors from copy-pasting tables and plots, but also ensures reproducibility of the data analysis \cite{baumer2014r-markdown}. $\mathsf{R}$ Markdown can generate reports in PDF, HTML, or Microsoft Word formats. $\mathsf{R}$ Markdown also provides a template for presentation slides, and the output can be in PDF, HTML, or Microsoft PowerPoint files. Note that PDF output from $\mathsf{R}$ Markdown requires the installation of LaTeX.^[<https://bookdown.org/yihui/rmarkdown/pdf-document.html>] (An alternative way is to save the HTML output to a PDF, but this is a manual step outside the end-to-end pipeline.) For a full book with multiple chapters, we recommend using the $\mathsf{R}$ package `bookdown` \cite{r-bookdown} for a comprehensive and reproducible workflow.  

The graphs in this manuscript are mainly for demonstration; they are created using the basic `plot` function as the default in RStudio. For advanced data visualization, we highly encourage readers to use the $\mathsf{R}$ package `ggplot2` \cite{ggplot2}. This package provides much more capability to convey the message in a graph, and the foundations behind `ggplot2` are based on *The Grammar of Graphics* \cite{wilkinson2012grammar}. For a simple and hands-on guide for data visualization, we recommend the book *Visualize This* \cite{yau2011visualize} to determine which types of graphs are most appropriate for which scenarios. On the other hand, *Semiology of Graphics: Diagrams, Networks, Maps* \cite{bertin1983semiology} is a comprehensive textbook for data visualization theory. Concepts such as human perception are technology-agnostic; the principles apply to almost any tool for data visualization.     

\section*{\textcolor{red}{Unfinished below}}  

The $\mathsf{R}$ code is also for demonstration. The author makes every effort to ensure correctness and readability of the code, but errors and inconsistencies are inevitable. Since this is an open-source project on GitHub,^[<https://github.com/star1327p/Exam-Scores-PTT>], readers are welcome to make a pull request and/or report any issues. To learn more about better coding styles, we recommend the computer science books *Clean Code* \cite{martin2009clean} and *Code Complete* \cite{mcconnell2004code}. For more complicated operations in $\mathsf{R}$, we recommend the $\mathsf{R}$ package `tidyverse` \cite{r-tidyverse} because it provides a fast, efficient, and organized workflow for general data modeling.^[<https://www.r-bloggers.com/2018/09/why-learn-the-tidyverse/>] The book \textit{$\mathsf{R}$ for Data Science: Import, Tidy, Transform, Visualize, and Model Data} \cite{wickham2016r} demonstrates an end-to-end cycle of using `tidyverse` in data science.  

\textcolor{red}{Talk about Python, `numpy`, `pandas`, and `scikit-learn`, etc?!}  

# Final: Personal Scores and Remarks {#personal-remarks}

```{r include-personal, child = '12-Personal-Scores-Remarks.Rmd'}
```

# Acknowledgments {.unnumbered}

The author would like to thank Dr. Mine Cetinkaya-Rundel and Dr. David Banks at Duke University; they both motivated the author to teach statistics and create reproducible work in $\mathsf{R}$. The author is also grateful to her Microsoft colleagues Smit Patel and Dylan Stout for troubleshooting GitHub issues.  

The author would also like to acknowledge Dr. Cliburn Chan and Dr. Janice McCarthy for introducing her to GitHub in the statistical computation course at Duke University. This provided her the foundations to use GitHub as a modern version control system in the first place.  

Finally, the author gives a special mention to her significant other, Hugh Hendrickson, for all his support in the author's professional career development.



# References {.unnumbered #references}