---
title: "PTT Analysis of Entrance Exam Scores in Taiwan"
subtitle: "A Statistical Approach with R Code"
author:
  - Christine P. Chai
  - Microsoft
  - cpchai21@gmail.com
date: \today
output: 
        pdf_document:
                number_sections: true
                citation_package: natbib
header-includes: 
  - \renewcommand{\and}{\\}
bibliography: references.bib
biblio-style: apalike
link-citations: yes
---

\renewcommand{\cite}{\citep}

```{r latex-cite-command, include=FALSE}
# %\let\cite\citep
# % from \citep to \cite to cite in author style, e.g. [Mule, 2008]

# % \bibliographystyle{plainnat}
# %\citep: citation in parentheses, e.g. [Mule, 2008]
# %\citet: citation as author, e.g. Mule [2008]
# %\cite: citation as author, \citet by default 
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Ongoing work since 2019.

# Executive Summary {.unnumbered}

\textcolor{red}{Write something here}

# Disclaimer {.unnumbered}

The opinions and views expressed in this manuscript are those of the author and do not necessarily state or reflect those of Microsoft. The author has a statistics and engineering background, without any training in middle school or high school teaching.

# Introduction {#intro}

```{r include-intro, child = '01-Introduction.Rmd'}
```

# Background {#background}

\section*{\textcolor{red}{Unfinished below}}

Elaborate on the following to enhance the background context:

- College entrance exam GSAT and AST: New rules in 2022 (not related to COVID)

- COVID-related rules for students who could not attend the exam  
Start writing about 2021 for both exams, because 2022 is still ongoing  

\textcolor{red}{This is the file we want to move content into.}  

\section*{\textcolor{red}{Unfinished above}}

```{r include-background, child = '02-Background.Rmd'}
```

# Data Description {#data}

\section*{\textcolor{red}{Unfinished below}}



Add the new PTT image and state **when** the registration restrictions were removed.  

Note that from 2018 to 2021, PTT limited new account registrations to only people with an email address from National Taiwan University.^[Screenshot obtained on May 26, 2020: <https://imgur.com/33fwrGH>]   

The good news is that this restriction has been removed.^[Screenshot obtained on February 5, 2022: <https://imgur.com/kaTsRpn>] As of 2022, anyone can register for an account on PTT with email^[<https://www.techbang.com/posts/88695-ptt-registration>] or mobile phone authentication^[<https://free.com.tw/ptt-id-registration-aotp/>].  

Also, list several mobile apps for readers to access PTT on a smart phone. Common apps include BePTT, Mo PTT, JPTT, and PiTT.  

\textcolor{red}{This is the file we want to move content into.}  

\section*{\textcolor{red}{Unfinished above}}

```{r include-data-description, child = '03-Data-Description.Rmd'}
```

# Exploratory Data Analysis {#eda}

```{r include-exploratory, child = '04-Exploratory.Rmd'}
```

# Linear Regression {#linear-reg}

```{r include-linear-reg, child = '05-Linear-Regression.Rmd'}
```

# Top Scorers: A Closer Look {#explore-top}

\section*{\textcolor{red}{Unfinished below}}  

Need to elaborate on the foreword in this chapter.  

Bivariate exploration subsection:  

Add my thoughts about categorical data analysis (e.g. proportional hazards model?!)  

\textcolor{red}{Add chi-squared test to quantify the level of significance}^[<https://data-flair.training/blogs/chi-square-test-in-r/>]      

What about ROC and AUC (area under the curve) with various probability thresholds?  

\textcolor{red}{This is the file we want to move content into.}  

\section*{\textcolor{red}{Unfinished below}}  

```{r include-linear-reg, child = '06-Top-Scorers.Rmd'}
```

\textcolor{red}{Should I create a new subsection for hypothesis testing and Pearson's chi-squared test?}  

We can perform hypothesis testing to statistically validate the positive association. The hypotheses are set up as follows:   

- $H_0$ (null hypothesis): No association exists between the categorical counts of **HighSchool_PR** and **College_Score**.  

- $H_1$ (alternative hypothesis): A positive association exists between the categorical counts of **HighSchool_PR** and **College_Score**.  

The p-value measures statistical significance, and it is the probability of observing something at least as extreme as the data, given the assumption that H0 is true \cite{diez2019openintro}.  

When p-value < 0.05, we reject $H_0$ and conclude that $H_1$ is true.  

When p-value > 0.05, we fail to reject $H_0$ and conclude that $H_0$ is true.  

(But note that we don't know the underlying truth and we can make mistakes. Should I talk about Type I error and Type II error?)  

The definition of p-value is often confused with the probability of the null hypothesis being true, which is incorrect \cite{goodman2008dirty}.   

\textcolor{red}{Consider adding a footnote/remark to link to Bayesian hypothesis testing and real posterior probabilities.}   

Bayesian hypothesis testing \cite{kruschke2018bayesian}  

P-values have been widely used and misused in scientific research, up to the extent that the \citet{american2016statement} released a statement on the proper use and interpretation of the p-values. Ideally we should consider the whole research design and the results' practical importance, rather than make conclusions solely based on statistical significance \cite{wasserstein2019moving}.  

Pearson's chi-squared test ($\chi^2$) can be used to evaluate whether the categorical counts of **HighSchool_PR** and **College_Score** are due to random chance. The test statistic $\chi^2$ is defined as below, and it asymptotically approaches a $\chi^2$ distribution.  

$$\chi^2 = \sum^{n}_{i=1}\dfrac{(O_i - E_i)^2}{E_i}$$

There are $n$ cells in the table. For each cell $i$, $O_i$ is the number of observations in that cell and $E_i$ is the number of expected counts under the null hypothesis.   

Then the test statistic $\chi^2$ is compared with a $\chi^2$ distribution to calculate the p-value, given the degrees of freedom. When we generate $m$ independent scores from a random sample, we have $m-1$ degrees of freedom because the $m$ scores are restricted by their sample mean. When we have $a$ rows and $b$ columns in the table, we have $(a-1)(b-1)$ degrees of freedom because each row and each column are restricted to the subtotal, on behalf of the grand total (total number of observations in the data).   

\textcolor{red}{Some foreword before using the chisq.test function to calculate}  

```{r chi-squared-test}
chisq.test(data_corr$HS_range, data_corr$CS_range)
```

\textcolor{red}{Write the result interpretation here}  

Why does this variable has 9 degrees of freedom?  

Since p-value < 0.05, what is our conclusion?   


\textcolor{red}{Also need to calculate the test statistic by hand}  

Goal is to verify that we are using the `chisq.test` function correctly.  
The readers should always leverage a package to do the computation when there is one that can do the job.    

Grand total = 188 records  
With 16 cells, $E_i = 188/16 = 11.75$  

Continue the calculation here, using for loops.  

# Logistic Regression {#logit-reg}

```{r include-logit-reg, child = '07-Logistic-Regression.Rmd'}
```

# Model Validation: In-Sample Prediction {#validation}

```{r include-in-sample-00, child = '08-In-Sample-00-Foreword.Rmd'}
```

## Implementation of In-Sample Prediction {#in-sample}

```{r include-in-sample-01, child = '08-In-Sample-01-Implementation.Rmd'}
```

## Interpretation of Confusion Matrix {#interpretation}

```{r include-in-sample-02, child = '08-In-Sample-02-Confusion-Matrix.Rmd'}
```

## Breakdown by High School Entrance Exam Scores {#another-breakdown}

```{r include-in-sample-03, child = '08-In-Sample-03-Score-Breakdown.Rmd'}
```

# Model Validation: Out-of-Sample Prediction {#out-of-sample}

```{r include-out-sample-00, child = '09-Out-Sample-00-Foreword.Rmd'}
```

## Separate Training and Testing Datasets

```{r include-out-sample-01, child = '09-Out-Sample-01-Sep-Explain.Rmd'}
```

### Implementation {#train-test-demo}

```{r include-out-sample-02, child = '09-Out-Sample-02-Sep-Implement.Rmd'}
```

### Organizing the Code for Reusability

```{r include-out-sample-03, child = '09-Out-Sample-03-Sep-Organize.Rmd'}
```

## Cross Validation

```{r include-out-sample-04, child = '09-Out-Sample-04-Cross-Validation.Rmd'}
```

### K-fold Cross Validation

```{r include-out-sample-05, child = '09-Out-Sample-05-K-Fold.Rmd'}
```

### Leave-one-out Cross Validation

```{r include-out-sample-06, child = '09-Out-Sample-06-Leave-One-Out.Rmd'}
```

## Comparison of Results {#cmp-results}

```{r include-out-sample-07, child = '09-Out-Sample-07-Comparison-Results.Rmd'}
```

# Recap of the Project {#recap}

```{r include-project-recap, child = '10-Project-Recap.Rmd'}
```

# Recommended Resources for Learning {#resources}

```{r include-additional-resources, child = '11-Additional-Resources.Rmd'}
```

# Final: Personal Scores and Remarks {#personal-remarks}

```{r include-personal, child = '12-Personal-Scores-Remarks.Rmd'}
```

# Acknowledgments {.unnumbered}

The author would like to thank Dr. Mine Cetinkaya-Rundel and Dr. David Banks at Duke University; they both motivated the author to teach statistics and create reproducible work in $\mathsf{R}$. The author is also grateful to her Microsoft colleagues Smit Patel and Dylan Stout for troubleshooting GitHub issues.  

The author would also like to acknowledge Dr. Cliburn Chan and Dr. Janice McCarthy for introducing her to GitHub in the statistical computation course at Duke University. This provided her the foundations to use GitHub as a modern version control system in the first place.  

Finally, the author gives a special mention to her significant other, Hugh Hendrickson, for all his support in the author's professional career development.



# References {.unnumbered #references}