---
title: "Statistical Analysis of High School and College Entrance Exam Scores in Taiwan with Online Data"
subtitle: "A Fully Reproducible Approach with R Markdown"
author:
  - Christine P. Chai
  - Microsoft
  - cpchai21@gmail.com
date: \today
output: 
        pdf_document:
                number_sections: true
                citation_package: natbib
header-includes: 
  - \renewcommand{\and}{\\}
bibliography: references.bib
biblio-style: apalike
link-citations: yes
---

\renewcommand{\cite}{\citep}

```{r latex-cite-command, include=FALSE}
# %\let\cite\citep
# % from \citep to \cite to cite in author style, e.g. [Mule, 2008]

# % \bibliographystyle{plainnat}
# %\citep: citation in parentheses, e.g. [Mule, 2008]
# %\citet: citation as author, e.g. Mule [2008]
# %\cite: citation as author, \citet by default 
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

def.chunk.hook  <- knitr::knit_hooks$get("chunk")
knitr::knit_hooks$set(chunk = function(x, options) {
  x <- def.chunk.hook(x, options)
  ifelse(options$size != "normalsize", paste0("\n \\", options$size,"\n\n", x, "\n\n \\normalsize"), x)
})
```

Ongoing work since 2019.

# Executive Summary {.unnumbered}

In this project, we investigate the relationship between high school and college entrance exam scores in Taiwan using self-reported data online. The goal is to demonstrate reproducible statistical analysis in $\mathsf{R}$ Markdown and LaTeX to PDF. We start by exploring the data to formulate the problem statement and identify the appropriate model, which can be an iterative process. We eventually decided on a binary outcome of the college entrance exams, and hence implemented a logistic regression to fit the model. We also validated the model with out-of-sample prediction methods, including cross validation as well as separate training and testing sets. The end-to-end data analysis can be compiled with a single button in RStudio, without copying-and-pasting output from one program to another.   


# Disclaimer {.unnumbered}

The opinions and views expressed in this manuscript are those of the author and do not necessarily state or reflect those of Microsoft. The author has a statistics and engineering background, without any training in middle school or high school teaching.

# Introduction {#intro}

```{r include-intro, child = '01-Introduction.Rmd'}
```

# Background {#background}

\section*{\textcolor{red}{Unfinished below}}

Elaborate on the following to enhance the background context:

- College entrance exam GSAT and AST: New rules in 2022 (not related to COVID)  

- COVID-related rules for students who could not attend the exam.    
Start writing about 2020, 2021, 2022 for both exams.   

- If we continue writing this manuscript in 2023, we absolutely need to add the recent changes in 2021 and 2022.  

\textcolor{red}{This is the file we want to move content into.}  

\section*{\textcolor{red}{Unfinished above}}

```{r include-background, child = '02-Background.Rmd'}
```

# Data Description {#data}

```{r include-data-description, child = '03-Data-Description.Rmd'}
```

# Exploratory Data Analysis {#eda}

```{r include-exploratory, child = '04-Exploratory.Rmd'}
```

# Linear Regression {#linear-reg}

```{r include-linear-reg, child = '05-Linear-Regression.Rmd'}
```

# Top Scorers: A Closer Look {#explore-top}

```{r include-linear-reg, child = '06-Top-Scorers-01-Univariate.Rmd'}
```

## Bivariate Exploration of Top Scorers {#bivariate-top-scorers}

```{r include-chisq-test, child = '06-Top-Scorers-02-Bivariate.Rmd'}
```

# Logistic Regression {#logit-reg}

```{r include-logit-reg, child = '07-Logistic-Regression.Rmd'}
```

# Model Validation: In-Sample Prediction {#validation}

```{r include-in-sample-00, child = '08-In-Sample-00-Foreword.Rmd'}
```

## Implementation of In-Sample Prediction {#in-sample}

```{r include-in-sample-01, child = '08-In-Sample-01-Implementation.Rmd'}
```

## Interpretation of Confusion Matrix {#interpretation}

```{r include-in-sample-02, child = '08-In-Sample-02-Confusion-Matrix.Rmd'}
```

## Breakdown by High School Entrance Exam Scores {#another-breakdown}

\section*{\textcolor{red}{Unfinished below}}  

Add my thoughts about categorical data analysis (e.g. proportional hazards model?!)  

Example: When **HighSchool_PR** is increased from Category X to Category Y, how does it affect the categories of **College_Score**?  

You don't have to do the full implementation, just write a few lines as a pointer to the concepts.  

\textcolor{red}{This is the file we want to move content into.}  

\section*{\textcolor{red}{Unfinished above}}   

```{r include-in-sample-03, child = '08-In-Sample-03-Score-Breakdown.Rmd'}
```

# Model Validation: Out-of-Sample Prediction {#out-of-sample}

\section*{\textcolor{red}{Unfinished below}}  

\textcolor{red}{Add a new section on ROC-AUC (likely to be Section 9.4)}   

\textcolor{red}{This is the file we want to move content into.}  

\section*{\textcolor{red}{Unfinished above}}  

```{r include-out-sample-00, child = '09-Out-Sample-00-Foreword.Rmd'}
```

## Separate Training and Testing Datasets {#sep-train-test}

```{r include-out-sample-01, child = '09-Out-Sample-01-Sep-Explain.Rmd'}
```

### Implementation {#train-test-demo}

```{r include-out-sample-02, child = '09-Out-Sample-02-Sep-Implement.Rmd'}
```

### Organizing the Code for Reusability {#org-code-reuse}

```{r include-out-sample-03, child = '09-Out-Sample-03-Sep-Organize.Rmd'}
```

## Cross Validation {#cross-validation}

```{r include-out-sample-04, child = '09-Out-Sample-04-Cross-Validation.Rmd'}
```

### K-fold Cross Validation {#k-fold}

```{r include-out-sample-05, child = '09-Out-Sample-05-K-Fold.Rmd'}
```

### Leave-one-out Cross Validation {#leave-one-out}

```{r include-out-sample-06, child = '09-Out-Sample-06-Leave-One-Out.Rmd'}
```

## Comparison of Results {#cmp-results}

```{r include-out-sample-07, child = '09-Out-Sample-07-Comparison-Results.Rmd'}
```

\section*{\textcolor{red}{Unfinished below}}  

# Model Metrics: ROC and AUC {#roc-auc}
```{r include-roc-auc, child = '10-Model-Metrics-ROC-AUC.Rmd'}
```

\section*{\textcolor{red}{Unfinished above}}  

# Recap of the Project {#recap}

```{r include-project-recap, child = '11-Project-Recap.Rmd'}
```

# Recommended Resources for Learning {#resources}

```{r include-additional-resources, child = '12-Additional-Resources.Rmd'}
```

# Final: Personal Scores and Remarks {#personal-remarks}

```{r include-personal, child = '13-Personal-Scores-Remarks.Rmd'}
```

# Acknowledgments {.unnumbered}

The author would like to thank Dr. Mine Cetinkaya-Rundel and Dr. David Banks at Duke University; they both motivated the author to teach statistics and create reproducible work in $\mathsf{R}$. The author is also grateful to her Microsoft colleagues Smit Patel and Dylan Stout for troubleshooting GitHub issues.  

The author would also like to acknowledge Dr. Cliburn Chan and Dr. Janice McCarthy for introducing her to GitHub in the statistical computation course at Duke University. This provided her the foundations to use GitHub as a modern version control system in the first place.  

Finally, the author gives a special mention to her significant other, Hugh Hendrickson, for all his support in the author's professional career development.



# References {.unnumbered #references}