---
title: "PTT Analysis of Entrance Exam Scores in Taiwan"
subtitle: "A Statistical Approach with R Code"
author:
  - Christine P. Chai
  - Microsoft
  - cpchai21@gmail.com
date: \today
output: 
        pdf_document:
                number_sections: true
                citation_package: natbib
header-includes: 
  - \renewcommand{\and}{\\}
bibliography: references.bib
biblio-style: apalike
link-citations: yes
---

\renewcommand{\cite}{\citep}

```{r latex-cite-command, include=FALSE}
# %\let\cite\citep
# % from \citep to \cite to cite in author style, e.g. [Mule, 2008]

# % \bibliographystyle{plainnat}
# %\citep: citation in parentheses, e.g. [Mule, 2008]
# %\citet: citation as author, e.g. Mule [2008]
# %\cite: citation as author, \citet by default 
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Ongoing work since 2019.

# Executive Summary {.unnumbered}

\textcolor{red}{Write something here}

## Disclaimer {.unnumbered}

The opinions and views expressed in this manuscript are those of the author and do not necessarily state or reflect those of Microsoft. The author has a statistics and engineering background, without any training in middle school or high school teaching.

# Introduction {#intro}

\section*{\textcolor{red}{Unfinished below}}

Consider making the introduction statistics-oriented, because we would like to focus on the statistical methodology. Most important is how to handle a real-time application problem.  

State this project as open-source and provide the GitHub link somewhere in the manuscript. Note that the name of GitHub repo may be changed later.  

Edit the paragraph of this manuscript's structure! Use cross-references for each chapter.  

\section*{\textcolor{red}{Unfinished above}}

```{r include-intro, child = '01-Introduction.Rmd'}
```

# Background {#background}

\section*{\textcolor{red}{Unfinished below}}

Elaborate on the following to enhance the background context:

- Starting in 2019, high school students are required to maintain a portfolio each semester to document their learning outcomes.^[<https://www.108epo.com/courses.php>]

- College entrance exam GSAT and AST: New rules in 2022 (not related to COVID)

- COVID-related rules for students who could not attend the exam  
Start writing about 2021 for both exams, because 2022 is still ongoing

\section*{\textcolor{red}{Unfinished above}}

```{r include-background, child = '02-Background.Rmd'}
```

# Data Description {#data}

\section*{\textcolor{red}{Unfinished below}}

Add the new PTT image and state **when** the registration restrictions were removed.  

Note that from 2018 to 2021, PTT limited new account registrations to only people with an email address from National Taiwan University.^[Screenshot obtained on May 26, 2020: <https://imgur.com/33fwrGH>]   

The good news is that this restriction has been removed.^[Screenshot obtained on February 5, 2022: <https://imgur.com/kaTsRpn>] As of 2022, anyone can register for an account on PTT with email^[<https://www.techbang.com/posts/88695-ptt-registration>] or mobile phone authentication^[<https://free.com.tw/ptt-id-registration-aotp/>].  

Also, list several mobile apps for readers to access PTT on a smart phone. Common apps include BePTT, Mo PTT, JPTT, and PiTT.

\section*{\textcolor{red}{Unfinished above}}

```{r include-data-description, child = '03-Data-Description.Rmd'}
```

# Exploratory Data Analysis {#eda}

```{r include-exploratory, child = '04-Exploratory.Rmd'}
```

# Linear Regression {#linear-reg}

\section*{\textcolor{red}{Unfinished below}}  

Elaborate on the foreword, because it cannot be just two lines.  

At least explain to readers what exactly is a linear regression.  

\section*{\textcolor{red}{Unfinished above}}  

```{r include-linear-reg, child = '05-Linear-Regression.Rmd'}
```

# Top Scorers: A Closer Look {#explore-top}

\section*{\textcolor{red}{Unfinished below}}  

Bivariate exploration subsection:  

Add my thoughts about categorical data analysis (e.g. proportional hazards model?!)

\section*{\textcolor{red}{Unfinished above}}  

```{r include-linear-reg, child = '06-Top-Scorers.Rmd'}
```

# Logistic Regression {#logit-reg}

```{r include-logit-reg, child = '07-Logistic-Regression.Rmd'}
```

# Model Validation: In-Sample Prediction {#validation}

```{r include-in-sample-00, child = '08-In-Sample-00-Foreword.Rmd'}
```

## Implementation of In-Sample Prediction {#in-sample}

```{r include-in-sample-01, child = '08-In-Sample-01-Implementation.Rmd'}
```

## Interpretation of Confusion Matrix {#interpretation}

```{r include-in-sample-02, child = '08-In-Sample-02-Confusion-Matrix.Rmd'}
```

## Breakdown by High School Entrance Exam Scores {#another-breakdown}

```{r include-in-sample-03, child = '08-In-Sample-03-Score-Breakdown.Rmd'}
```

# Model Validation: Out-of-Sample Prediction {#out-of-sample}

```{r include-out-sample-00, child = '09-Out-Sample-00-Foreword.Rmd'}
```

## Separate Training and Testing Datasets

```{r include-out-sample-01, child = '09-Out-Sample-01-Sep-Explain.Rmd'}
```

### Implementation {#train-test-demo}

```{r include-out-sample-02, child = '09-Out-Sample-02-Sep-Implement.Rmd'}
```

### Organizing the Code for Reusability

```{r include-out-sample-03, child = '09-Out-Sample-03-Sep-Organize.Rmd'}
```

## Cross Validation

```{r include-out-sample-04, child = '09-Out-Sample-04-Cross-Validation.Rmd'}
```

### K-fold Cross Validation

```{r include-out-sample-05, child = '09-Out-Sample-05-K-Fold.Rmd'}
```

### Leave-one-out Cross Validation

```{r include-out-sample-06, child = '09-Out-Sample-06-Leave-One-Out.Rmd'}
```

## Comparison of Results {#cmp-results}

```{r include-out-sample-07, child = '09-Out-Sample-07-Comparison-Results.Rmd'}
```

# Recap of the Project {#recap}

```{r include-project-recap, child = '10-Project-Recap.Rmd'}
```

# Additional Resources {#resources}

```{r include-additional-resources, child = '11-Additional-Resources.Rmd'}
```

# Final: Personal Remarks

\section*{\textcolor{red}{Unfinished below}}

\textcolor{red}{Continue to separate this enormous \texttt{.Rmd} into smaller files.}  

Finally, include my own scores as a datapoint for prediction.  

The author scored **HighSchool_PR** 95 in Year 2004, but she got admitted to Taipei First Girls' High School because the special class for math and science track had a separate admission process at that time. Taipei First Girls' High School^[http://web.fg.tp.edu.tw/~tfghweb/EnglishPage/index.php] typically requires **HighSchool_PR** 99 for admission, with few exceptions such as recruited athletes^[<https://www.sa.gov.tw/PageContent?n=1265>] and students with disabilities^[<http://www.rootlaw.com.tw/LawArticle.aspx?LawID=A040080080001900-1020822>]. Several years later, the policy changed so that students had to be admitted to the high school first in order to try for the special class placement within. Most students in the special class had **HighSchool_PR** 99 anyway, and the author was one of the few who did not. Hence the author wondered whether she would have similar academic performance, had she enrolled in a different high school commensurate to **HighSchool_PR** 95 in Taipei.    


The author got **College_Score** 69 on the General Scholastic Ability Test (GSAT) in Year 2007. Later that year, she took the Advanced Subjects Test (AST) and got admitted to The Department of Electrical Engineering at National Taiwan University (NTUEE).   

State the admissions and rejections^[<https://sites.google.com/site/christinepeijinnchai/resume-cv>]  

Write something more  

NTUEE^[https://web.ee.ntu.edu.tw/eng/index.php] typically requires
full marks (15 out of 15) in English, mathematics, and science in **College_Score** for the early admission through GSAT.^[https://university.1111.com.tw/univ_depinfo9.aspx?sno=100102&mno=520101] Most students at NTUEE had a **College_Score** of 70 or higher, at the time when 75 was the max possible score. But still a significant number of students got admitted through the AST in July.  

Treat this part as an anecdote and explain my motivation.  

Instead of dwelling on the past, the author decided to redirect her energy to something more meaningful and constructive. That's why this manuscript was born.      

In addition to statistical analysis, she also wanted to demonstrate full reproducibility of the project.  


Inspired by \citet{baumer2014r-markdown} $\Rightarrow$ $\mathsf{R}$ Markdown   

Write something more


# Acknowledgments {.unnumbered}

The author would like to thank Dr. Mine Cetinkaya-Rundel and Dr. David Banks at Duke University; they both motivated the author to teach statistics and create reproducible work in `R`. The author is also grateful to her Microsoft colleagues Smit Patel and Dylan Stout for troubleshooting GitHub issues.  

Finally, the author would like to acknowledge Dr. Cliburn Chan and Dr. Janice McCarthy for introducing her to GitHub in the statistical computation course at Duke University. This provided her the foundations to use GitHub as a modern version control system in the first place.

# Appendix {.unnumbered}

\section*{\textcolor{red}{Unfinished below}}

```{r include-appendix, child = 'Appendix.Rmd'}
```


# References {.unnumbered #references}